Tight Bounds for Dynamic Balanced Graph Partitioning

### 2020-05-09 paper/disc20clustering.tex
On k=3

\begin{figure}[H]
	\centering
	\includegraphics[width=0.3\textwidth]{figs/substitute}
	\caption{an unsaturated edge that is inside a component}
\end{figure}

\begin{theorem}
	\TAlg is $O(\ell)$-competitive for $k=3$.
\end{theorem}
\begin{proof}
	Fix a completed phase, and consider the state of \TAlg's counters at the end of it.
	We consider the incomplete phase later in this proof.

@@

	\medskip
	Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to requests $\sigma_E$.
	By entering some fixed configuration before serving $\sigma_E$, \OPT may mitigate paying for requests between at most $3\ell$ pairs of nodes by collocating them in its clusters.
	Recall that $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for requests that increased the counter (i.e., external requests), thus $\OPT$ may mitigate at most $3\ell\cdot(\alpha - 1)$ requests from $\sigma_E$.
	Faced with $W := |\sigma_E| - 3\ell\cdot(\alpha-1)$ requests $\sigma_E$ that it could not mitigate, $\OPT$ served some of them remotely and possibly performed some migrations to decrease its cost.

	Now we estimate the cost of $\OPT(\sigma_E)$ while accounting savings from migrations.
	By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
	This may allow serving requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
	As $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
	By performing a single swap that costs $2\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$
	Thus, for serving $\sigma_E$, $\OPT$ pays at least $\OPT(\sigma_E) + \OPT(M) \geq W \cdot \frac{2\alpha}{4 (\alpha-1)}\geq |\sigma_E| / 2 - 2 \alpha \cdot \ell$.
	From that, we obtain $|\sigma_E| \geq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
	Let $E := \OPT(\sigma_E) + \OPT(M)$.

	Finally, we bound the competitive ratio
	\begin{equation*}
		\frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{11\alpha \cdot \ell + |\sigma_E|}{\alpha/2 + E/2} \leq \frac{27\alpha\cdot\ell + 2\cdot E}{\alpha + E} \leq 27 \ell = O(\ell).
	\end{equation*}

@@

	It remains to consider the last, unfinished phase.
	First, consider the case, where the unfinished phase is also the first one.
	Then, we cannot charge $\OPT$ due to the inability to partition the components.
	Instead, we use the fact that \TAlg and $\OPT$ started with the same initial configuration.
	If the input finished before the first $\alpha$ external requests, then \TAlg is $1$-competitive.
	If there were at least $\alpha$ external requests, then $\OPT$ either paid $\alpha$ for serving them remotely, or paid $\alpha$ for a migration; for the cost of \TAlg we follow the analysis of the unsaturated pairs.
	Second, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.
	This way, the competitive ratio increases at most twofold.
\end{proof}

>>>>

\begin{figure}[H]
	\centering
	\includegraphics[width=0.3\textwidth]{figs/substitute}
	\caption{an unsaturated edge that is inside a component. If there is a space next to it, maybe cluster types should fit next to it? Not sure.}
\end{figure}

\begin{theorem}
	\TAlg is $(54\ell)$-competitive for $k=3$.
\end{theorem}
\begin{proof}
	Fix a completed phase, and consider the state of \TAlg's counters at the end of it.
	We consider the incomplete phase later in this proof.

@@

	\medskip
	Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to requests $\sigma_E$.
	By entering some fixed configuration before serving $\sigma_E$, \OPT may mitigate paying for requests between at most $3\ell$ pairs of nodes by collocating them in its clusters.
	Recall that $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for requests that increased the counter (i.e., external requests), thus $\OPT$ may mitigate at most $3\ell\cdot(\alpha - 1)$ requests from $\sigma_E$.
	Faced with at least $W := |\sigma_E| - 3\ell\cdot(\alpha-1)$ requests $\sigma_E$ that it could not mitigate, $\OPT$ served some of them remotely and possibly performed some migrations to decrease its cost.

	Now we estimate the cost of $\OPT(\sigma_E)$ while accounting savings from migrations.
	By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
	This may allow serving requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
	As $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
	By performing a single swap that costs $2\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$
	Hence, for serving $\sigma_E$, $\OPT$ pays at least $\OPT(\sigma_E) + \OPT(M) \geq W \cdot \frac{2\alpha}{4 (\alpha-1)}\geq |\sigma_E| / 2 - 2 \alpha \cdot \ell$.

	Finally, we bound the competitive ratio.
	From the bound on \OPT's cost we have $|\sigma_E| \geq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
	Let $\xi := \OPT(\sigma_E) + \OPT(M)$.
	The competitive ratio is

	\[
		\frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{11\alpha \cdot \ell + |\sigma_E|}{\alpha/2 + \xi/2} \leq \frac{27\alpha\cdot\ell + 2\cdot \xi}{\alpha + \xi} \leq 27 \ell.
	\]

@@

	It remains to consider the last, unfinished phase.
	First, consider the case, where the unfinished phase is also the first one.
	Then, we cannot charge $\OPT$ due to the inability to partition the components.
	Instead, we use the fact that \TAlg and $\OPT$ started with the same initial configuration.
	If the input finished before the first $\alpha$ external requests, then \TAlg is $1$-competitive.
	If there were at least $\alpha$ external requests, then $\OPT$ either paid $\alpha$ for serving them remotely, or paid $\alpha$ for a migration; for the cost of \TAlg we follow the analysis of the unsaturated pairs.
	Second, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.
	This way, the competitive ratio increases at most twofold in comparison to a finished phase, and the competitive ratio is $\TAlg(\sigma)/\OPT(\sigma) \leq 54\ell$.
\end{proof}
